---
title: "Multigroup confirmatory factor analysis"
author: "Nikhil Agrawal"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

# Load packags
```{r}
library(lavaan)
library(MASS)
library(psych)
library(sirt)
```

# Load data
```{r}
data <- read.csv("Final_COVIDiSTRESS_Vol2_cleaned.csv")
# summary(data)
```

# Data preparation
## General

```{r}
data <- data[order(data$UserLanguage), ]
languages <- labels(table(data$UserLanguage))[[1]]
print(length(languages))

# Remove languages whose count is less than 100
data <- data[data$UserLanguage %in% languages[which(table(data$UserLanguage) > 100)], ]
languages <- labels(table(data$UserLanguage))[[1]]
print(length(languages))
```

## Vaccine attitudes

```{r}
data$vaccine_attitudes_4_midneutral <- 8 - data$vaccine_attitudes_4_midneutral
data$vaccine_attitudes_5_midneutral <- 8 - data$vaccine_attitudes_5_midneutral

va_vars <- c(
    "vaccine_attitudes_9_midneutral",
    "vaccine_attitudes_2_midneutral",
    "vaccine_attitudes_3_midneutral", "vaccine_attitudes_5_midneutral",
    "vaccine_attitudes_6_midneutral"
)

consp_vars <- c(
    "conspirational_think_1", "conspirational_think_2",
    "conspirational_think_3", "conspirational_think_4"
)

anti_vars <- c('antiexpert_1','antiexpert_2','antiexpert_3')
```


# Confirmatory factor analysis
## Choose measures
We choose the traditional criteria for configural level of invariance, which is
RMSEA $\leq$ .08, SRMR $\leq$ .08, CFI $\geq$ .90, TLI $\geq$ .90
```{r}
fits <- c("rmsea.scaled", "srmr", "cfi.scaled", "tli.scaled")
```

## Vaccine attitudes

```{r}
cfa.eq.va <- 'VA =~ vaccine_attitudes_9_midneutral + 
vaccine_attitudes_2_midneutral + vaccine_attitudes_3_midneutral+
vaccine_attitudes_5_midneutral + vaccine_attitudes_6_midneutral'

cfa.va.config <- cfa(cfa.eq.va,
    data = data,
    group = "UserLanguage", estimator = "WLSMV"
)
fitmeasures(cfa.va.config)[fits]
```
The data above doesn't fit our configural invariance criteria.
We check for metric invariance

```{r}
cfa.a.metric <- cfa(cfa.eq.va,
    data = data, group = "UserLanguage",
    estimator = "WLSMV", group.equal = "loadings"
)
fitmeasures(cfa.a.metric)[fits]

fitmeasures(cfa.a.metric)[fits] - fitmeasures(cfa.va.config)[fits]
```

The data above doesn't fit our metric invariance criteria either. Thus we need to align our data.

### Aligning the data
```{r results=FALSE, message=FALSE, warning=FALSE}
# Extract parameters
par.va <- invariance_alignment_cfa_config(
    dat = data[, va_vars], group = data$UserLanguage
)
```
```{r}
# Alignment
mod1.va <- invariance.alignment(
    lambda = par.va$lambda,
    nu = par.va$nu,
    align.scale = c(0.2, 0.4), align.pow = c(0.25, 0.25)
)

mod1.va$es.invariance['R2',]
```

## Conspirational Thinking

```{r}
cfa.eq.consp <- 'CONSP =~ conspirational_think_1 + conspirational_think_2+
  conspirational_think_3 + conspirational_think_4'

cfa.consp.config <- cfa(cfa.eq.consp,
    data = data,
    group = "UserLanguage", estimator = "WLSMV"
)
fitmeasures(cfa.consp.config)[fits]
```

```{r}
cfa.consp.metric <- cfa(cfa.eq.consp,
    data = data, group = "UserLanguage",
    estimator = "WLSMV", group.equal = "loadings"
)
fitmeasures(cfa.consp.metric)[fits]

fitmeasures(cfa.consp.metric)[fits] - fitmeasures(cfa.consp.config)[fits]
```

```{r results=FALSE, message=FALSE, warning=FALSE}
# Extract parameters
par.consp <- invariance_alignment_cfa_config(
    dat = data[, consp_vars], group = data$UserLanguage
)
```

```{r}
# Alignment
mod1.consp <- invariance.alignment(
    lambda = par.consp$lambda,
    nu = par.consp$nu,
    align.scale = c(0.2, 0.4), align.pow = c(0.25, 0.25)
)

mod1.consp$es.invariance['R2',]
```

## Anti expert attitudes
```{r}
cfa.eq.anti = "ANTI =~ antiexpert_1 + antiexpert_2 + antiexpert_3"

cfa.anti.config <- cfa(cfa.eq.anti,
    data = data,
    group = "UserLanguage", estimator = "WLSMV"
)

fitmeasures(cfa.anti.config)[fits]
```

```{r}
cfa.anti.metric <- cfa(cfa.eq.anti,
    data = data, group = "UserLanguage",
    estimator = "WLSMV", group.equal = "loadings"
)
fitmeasures(cfa.anti.metric)[fits]

fitmeasures(cfa.anti.metric)[fits] - fitmeasures(cfa.anti.config)[fits]
```

```{r results=FALSE, message=FALSE, warning=FALSE}
# Extract parameters
par.anti <- invariance_alignment_cfa_config(
    dat = data[, anti_vars], group = data$UserLanguage
)
```

```{r}
# Alignment
mod1.anti <- invariance.alignment(
    lambda = par.anti$lambda,
    nu = par.anti$nu,
    align.scale = c(0.2, 0.4), align.pow = c(0.25, 0.25)
)

mod1.anti$es.invariance['R2',]
```

# Factor score calculation

```{r}
aligned.factor.scores <- function(lambda,nu,y){
  lambda1 <- ginv((lambda))
  ns <- nrow(y)
  nus <- matrix(nu,nrow=ns,ncol=length(nu),byrow=T)
  y_nu <- y - nus
  F <- lambda1 %*% t(as.matrix(y_nu))
}


# Iterate over languages
for (i in 1:length(languages)) {
    # Select data
    lang <- languages[i]
    data_lang <- data[data$UserLanguage == lang, ]
    
    ns.va <- aligned.factor.scores(
        mod1.va$lambda.aligned[i, ], mod1.va$nu.aligned[i, ],
        data_lang[, va_vars]
    )
    
    ns.consp <- aligned.factor.scores(
        mod1.consp$lambda.aligned[i, ], mod1.consp$nu.aligned[i, ],
        data_lang[, consp_vars]
    )

    ns.anti <- aligned.factor.scores(
        mod1.anti$lambda.aligned[i, ], mod1.anti$nu.aligned[i, ],
        data_lang[, anti_vars]
    )

    # Calculate factor scores
    data_lang$vx <- t(ns.va)
    data_lang$consp <- t(ns.consp)
    data_lang$anti <- t(ns.anti)
    
    # Save data
    if (i == 1) {
        data_aligned <- data_lang
    } else {
        data_aligned <- rbind(data_aligned, data_lang)
    }
}
```

# Data filteration

```{r}
# Remove countries with less than 30 respondents

data_aligned <- data_aligned[order(data_aligned$residing_country), ]
countries <- labels(table(data_aligned$residing_country))[[1]]
print(length(countries))

data_filtered <- data_aligned[data_aligned$residing_country %in% countries[which(table(data_aligned$residing_country) > 100)], ]
countries <- labels(table(data_filtered$residing_country))[[1]]
print(length(countries))
```

```{r}
summary(data_filtered)
save(data_filtered, file = "data_filtered.RData")
```